---
title: "powercalcs by simulation"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

Libraries and variable

```{r setup, results='hide', message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(tibble)
library(rio)
library(reshape)
library(ggplot2)
library(limma)
library(truncnorm)
setwd("./") # set to directory with the code, metadata and data
```

Useful custom function

```{r functions, results='hide'}
do_de <- function(sim_ma, group){
    d = model.matrix(~ group)
    # y = normalizeBetweenArrays(sim_ma, method = "quantile")
    y = normalizeBetweenArrays(sim_ma, method = "cyclicloess")
    f = lmFit(y, d)
    fb = eBayes(f, robust = TRUE)
    topTable(fb,coef = "grouptreatment", sort.by = "none", number = Inf)
}
```

Do some data munging

```{r dataimport}
dt = import("data.xlsx") %>% janitor::clean_names() %>% as_tibble()
dt <- column_to_rownames(dt, "sample_id")
dt <- t(dt)

meta <- import("meta.xlsx")

# merge metadata into the samplenames
all(colnames(dt)==meta$sample_id)
meta$newnames <- paste( meta$group,meta$sample_id, sep="_")
colnames(dt) <- meta$newnames
```

Setup simulation comparisons
- here we we will do all possible pairwise comparisons between the 3 groups
- we also set the parameters to test in the simulation ie. fold change and sample number


```{r variables,results='hide'}
sampleclasses <- meta %>%  pull(group) %>% unique()
combos <-  combn(sampleclasses,2) %>% t() %>% as_tibble() 
# simulate 100x: abs(fc)=1, 1.5, 2, 4 with 5,10,25,50 people each
sim_fc = c(1, 1.5, 2, 4) # fold changes we are testing in the simulations
sim_n = c(2,3,4,5,6, 10) # number of samples we will test in the simulations
```

This is where we run the simulations


```{r simulations, echo=FALSE, results='hide'}
temp <- apply(combos, 1, function(combo){
  class1 <- paste(combo[1], "_", sep="")
  class2 <- paste(combo[2], "_", sep="")
  combo_title <- paste(combo[1], "vs", combo[2],  sep="_")
  
  ma = as.data.frame(dt[,c(grep(class1, colnames(dt)), grep(class2, colnames(dt)))])
  
  # get variance for each metabolite
  sd_v = apply(ma,1, sd)
  m_v = apply(ma,1, mean)
  # plot(m_v,sd_v)
  
  # example how to use it rtruncnorm(n=10, a=0, b=340, mean=39.4, sd=25.09)
  set.seed(4242)
  sims = expand.grid(a = sim_fc, b = sim_n)
  res_sim = apply(sims ,1,function(sim){
    # print(i)
    sim <- unlist(sim)
    lapply(1:100, function(j){
      # print(j)
      # pick some random metabolites to be the ones that become DE
      .de = sample(1:nrow(ma), 4)
      .up = .de[1:2]
      .down = .de[3:4]
      
      # rtruncnorm will generate random values based on the standard dev and mean 
       # we set n to be high enough to gie us 2 columns for each n, and the same number of metabolites as the original dataset
      # then we shove it in a matrix whose column numbers are based on the replicate size
      df = matrix(rtruncnorm(n=as.numeric(sim[2]*2*nrow(ma)),  
                             a=0, b=Inf, 
                             mean = m_v, 
                             sd = sd_v), 
                  ncol=as.numeric(sim[2]*2))
      
      # here we artificially force the randomly chosen metabolites to have fold changes
      df[.up, 1:sim[2]] = df[.up, 1:sim[2]] * sim[1]
      df[.down, 1:sim[2]] = df[.down, 1:sim[2]] / sim[1]
      

      group = c(rep("control", sim[2]), rep("treatment", sim[2]))
      res <- do_de(df, group)
      # print(res)
      .uprigh = sum(res[.up,"adj.P.Val"] < 0.05)
      .downrigh = sum(res[.down,"adj.P.Val"] < 0.05)
      .total = sum(res[,"adj.P.Val"] < 0.05)
      TP = .uprigh + .downrigh
      FP = .total - TP
      data.frame(tp=TP, fp=FP, n=sim[2], fc=sim[1], sim=j)
    }) %>% bind_rows()
  }) %>% bind_rows()
  
  dfp = res_sim %>% group_by(fc,n) %>%
    summarise(fpm=mean(fp),tpm=mean(tp), fpsd=sd(fp), tpsd=sd(tp), precision=tpm/(tpm+fpm))
  
  needed_replicates <- filter(dfp, precision>0.8) %>% group_by(fc) %>% summarise(min_replicates=min(n))
  
  return(list(combo=combo_title, dfp=dfp, reps=needed_replicates))
})
```

# Precisions plots

```{r plots}

plot_results <- function(results) {
  mytitle <- results$combo
  results$dfp %>% filter(fc!=1) %>%
    ggplot(aes(x=as.factor(n*2),y=tpm/(tpm+fpm), group=as.factor(fc), color=as.factor(fc))) +
    geom_line(stat = "identity", size=2) +
    scale_color_brewer("Fold change", palette = "Set1") +
    ylim(0,1) +
    xlab("Number samples") + ylab("PPV") +
    ggtitle(paste(mytitle, "Precision for lipids at 5% FDR", sep=" - "))
}
lapply(temp, plot_results)  
```


# Precision tables

```{r tables, results='asis'}
table_dfpresults <- function(results){
  dfp <- filter(results$dfp, fc!=1)  
  
  knitr::kable(dfp, caption=results$combo)
        # , caption=results$combo))
  # , extensions = 'FixedHeader',
  # options = list(pageLength = 10, fixedHeader = TRUE))
}

lapply(temp, table_dfpresults)
```

# Precision summaries
- number of replicates needed for 80% precision at the stated fold change

```{r finaltables, results='asis'}
table_finalresults <- function(results){

  knitr::kable(results$reps, caption=results$combo)
        # , caption=results$combo))
  # , extensions = 'FixedHeader',
  # options = list(pageLength = 10, fixedHeader = TRUE))
}

lapply(temp, table_finalresults)
```

